### 2014-02-05  feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/empirical.tex
Latest submitted version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[tbp!]
%\vspace{-1mm}
\centering
{\small 
\begin{tabular}{|l|r|r|r|r|l|} \hline
Dataset & \# Features & \# Data & \# Features/\# Data & \% True Labels & Feature Type\\ \hline \hline 
Breast Cancer(BC)	& 9	& 699	& 0.013	& 34 \%	& All Numeric\\ \hline
Diabetes(D)        & 8	& 768	& 0.010	& 35 \%	& All Numeric\\ \hline
Heart Statlog(HS)	& 13	& 270	& 0.048	& 44 \%	& All Numeric\\ \hline
Spect(S)	        & 22	& 80	& 0.275 & 33 \%	& Categorical\\ \hline
Vote(V)	        & 16	& 435	& 0.037	& 39 \%	& Categorical\\ \hline
Newsgroup(NG)	& 500	& 1963	& 0.255	& 49 \%	& All Binary\\ \hline
Horse Colic(HC)	& 22	& 368	& 0.060	& 37 \%	& Categorical (16) and numeric (6)\\ \hline
Credit-American(CR)	& 15	& 690	& 0.022	& 44 \%	& Categorical(9) and numeric (6)\\ \hline
Credit-German(GC)	& 20	& 1000	& 0.020 & 30 \%	& Categorical (13) and Numeric (7)\\ \hline
Hepatitis(H)	& 19	& 155	& 0.123	& 21 \%	& Categorical (13) and Numeric (6)\\ \hline
Ionosphere(I)	& 34	& 351	& 0.097	& 36 \%	& All Numeric\\ \hline
%KR-vs-KP	& 36	& 3196	& 0.011	& 48 \%	& All Categorical\\ \hline
%Labor	        & 16	& 57	& 0.281	& 35 \%	& Categorical (8) and Numeric (8)\\ \hline
%Mushroom	& 22	& 8124	& 0.003	& 48 \%	& All Categorical\\ \hline
%Sick	        & 29	& 3772	& 0.008	&  6 \%	& Categorical (23) and Numeric (6)\\ \hline
%Sonar	        & 60	& 208	& 0.288 & 47 \%	& All Numeric\\ \hline
\end{tabular}}
\vspace{-1mm}
\caption{\footnotesize Descriptive statistics of the various datasets evaluated in this work.}
\label{table:stats}
\vspace{-4mm}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\includegraphics[width=0.85\textwidth]{./figures/dataset_perf/Accuracy.pdf}\\
\includegraphics[width=0.85\textwidth]{./figures/dataset_perf/Recall.pdf}
\includegraphics[width=0.10\textwidth]{./figures/dataset_perf/legend.PNG}
\vspace{-2mm}
\caption{\footnotesize Performance of various feature selection algorithms per dataset, averaged across classifier type and each stage of feature selection.  Overall the newly proposed CVL method performs well on Recall on most datasets, while the Recall performance of other feature selection algorithms vary; interestingly the best performance is achieved for datasets a high ratio of features to data indicating that the strong conjunctive voting scheme of CVL may help reduce noise in feature selection for these datasets. }
\label{fig:perf_vs_dataset}
\vspace{-4mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\centering
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectLogisticRegressionRecall.pdf}
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectNaiveBayesRecall.pdf}\\
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupLogisticRegressionRecall.pdf}
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupNaiveBayesRecall.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectLogisticRegressionRecall2.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectNaiveBayesRecall2.pdf}\\
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupLogisticRegressionRecall2.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupNaiveBayesRecall2.pdf}
\caption{\footnotesize Performance distribution of various feature selection algorithms for two datasets and classifier types at each stage of feature selection.  The conjunctive voting methods do best overall in these datasets.}
\label{fig:perf_vs_fs_alg}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Empirical}

In this section, we conduct experiments to study the feature selection
performances of several well known methods such as Symmetrical
Uncertainty Rank (SUR), Gain Ratio Rank (GRR), Mutual Information Rank
(MIR), Correlation Based Rank (CR), Conditional Entropy Rank (CER)~\cite{guyon_jmlr03}, Correlation-based Feature Subset Selection (CS)~\cite{Hall1998}, Reliff (RF)~\cite{Robnik-Sikonja2003} , Minimum Redundancy Maximum Relevancem(MRMR)~\cite{peng2005} and our proposed methods:  Disjunction Voting Expectation(DVE) and  Disjunction Voting likelihood (DVL), Conjunction Voting Expectation (CVE), Conjunction Voting likelihood (CVL). The first five methods are variable rank methods and the last seven are subset rank methods. The first five methods are variable rank methods that select variables by ranking them with some metric and the last seven are subset rank methods that assess subsets of variables and consider previous selections to decide the next selection.

@@

In our evaluation, we wish to answer the following questions:
Do different feature selection methods perform better with
each classifier or on each dataset in comparison to others?
How reliably do the different feature selection methods perform overall
in terms of their performance distribution?
%\item Do any dataset features seem to correlate with algorithm performance?

To answer these questions, we first start with
Figure~\ref{fig:perf_vs_dataset}, where we evaluate the performance of
various feature selection algorithms per dataset, averaged across
classifier type and each stage of feature selection. We can observe that our approaches improve recall and accuracy in many of the binary classification problem.
% TODO Scott

In Figure~\ref{fig:perf_vs_classifier}, we examine how well our 
novel feature selection algorithms DVE, DVL, CVE, CVL 
perform vs. classifier, noting that overall performance is best
for Naive Bayes followed by Logistic Regression and noticeably worse for
the SVM Linear --- it seems simply that the probabilistic nature of our feature selection
dovetail best with classification methods that are themselves probabilistic,
and overall best with Naive Bayes which makes the same independence assumptions.

Next we examine Figure~\ref{fig:perf_vs_fs_alg}, where we evaluate the
performance distribution of various feature selection algorithms with
samples taken for each dataset, classifier type, and stage of feature
selection. This graph shows that our CVL high recall feature selection approach 
improves recall with the most improvement noticeable when there is a high
feature to data ratio such as in the Newsgroup (NG) and Spect (S) data
sets.  We conjecture the reason for this stems from the strong agreement of
features provided by the conjunctive voting scheme that helps reduce noise
in the feature selection process.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[tbp!]
%\vspace{-1mm}
\centering
{\small 
\begin{tabular}{|l|r|r|r|r|l|} \hline
Dataset & \# Features & \# Data & \# Features/\# Data & \% True Labels & Feature Type\\ \hline \hline 
Breast Cancer (BC)	& 9	& 699	& 0.013	& 34 \%	& All Numeric\\ \hline
Diabetes (D)        & 8	& 768	& 0.010	& 35 \%	& All Numeric\\ \hline
Heart Statlog (HS)	& 13	& 270	& 0.048	& 44 \%	& All Numeric\\ \hline
Spect (S)	        & 22	& 80	& 0.275 & 33 \%	& Categorical\\ \hline
Vote (V)	        & 16	& 435	& 0.037	& 39 \%	& Categorical\\ \hline
Newsgroup (NG)	& 500	& 1963	& 0.255	& 49 \%	& All Binary\\ \hline
Horse Colic (HC)	& 22	& 368	& 0.060	& 37 \%	& Categorical (16) and Numeric (6)\\ \hline
Credit-American (CR)	& 15	& 690	& 0.022	& 44 \%	& Categorical (9) and Numeric (6)\\ \hline
Credit-German (GC)	& 20	& 1000	& 0.020 & 30 \%	& Categorical (13) and Numeric (7)\\ \hline
Hepatitis (H)	& 19	& 155	& 0.123	& 21 \%	& Categorical (13) and Numeric (6)\\ \hline
Ionosphere (I)	& 34	& 351	& 0.097	& 36 \%	& All Numeric\\ \hline
%KR-vs-KP	& 36	& 3196	& 0.011	& 48 \%	& All Categorical\\ \hline
%Labor	        & 16	& 57	& 0.281	& 35 \%	& Categorical (8) and Numeric (8)\\ \hline
%Mushroom	& 22	& 8124	& 0.003	& 48 \%	& All Categorical\\ \hline
%Sick	        & 29	& 3772	& 0.008	&  6 \%	& Categorical (23) and Numeric (6)\\ \hline
%Sonar	        & 60	& 208	& 0.288 & 47 \%	& All Numeric\\ \hline
\end{tabular}}
\vspace{-2mm}
\caption{\footnotesize Descriptive statistics of the various datasets evaluated in this work.}
\label{table:stats}
\vspace{-3.5mm}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\includegraphics[width=0.85\textwidth]{./figures/dataset_perf/Accuracy.pdf}\\
\includegraphics[width=0.85\textwidth]{./figures/dataset_perf/Recall.pdf}
\includegraphics[width=0.10\textwidth]{./figures/dataset_perf/legend.PNG}
\vspace{-2mm}
\caption{\footnotesize Performance of various feature selection algorithms per dataset, averaged across classifier type and each stage of feature selection.  Overall the newly proposed CVL method performs well on Recall on most datasets, while the Recall performance of other feature selection algorithms vary; interestingly the best performance is achieved for datasets a high ratio of features to data indicating that the strong conjunctive voting scheme of CVL may help reduce noise in feature selection for these datasets. }
\label{fig:perf_vs_dataset}
\vspace{-3mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\centering
\includegraphics[width=0.38\textwidth]{./figures/linegraphs/SpectLogisticRegressionRecall.pdf}
\includegraphics[width=0.38\textwidth]{./figures/linegraphs/SpectNaiveBayesRecall.pdf}\\
\includegraphics[width=0.38\textwidth]{./figures/linegraphs/NewsgroupLogisticRegressionRecall.pdf}
\includegraphics[width=0.38\textwidth]{./figures/linegraphs/NewsgroupNaiveBayesRecall.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectLogisticRegressionRecall2.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectNaiveBayesRecall2.pdf}\\
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupLogisticRegressionRecall2.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupNaiveBayesRecall2.pdf}
\caption{\footnotesize Recall performance of various feature selection algorithms for two datasets and classifier types at each stage of feature selection.  The conjunctive voting methods (CVE and CVL)
perform well (and reliably) across all four combinations of classifier and dataset.}
\label{fig:perf_vs_fs_alg}
%\vspace{-2mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Empirical Evaluation}

In this section, we conduct experiments to study the feature selection
performances of several well known methods such as Symmetrical
Uncertainty Rank (SUR), Gain Ratio Rank (GRR), Mutual Information Rank
(MIR), Correlation Based Rank (CR), Conditional Entropy Rank
(CER)~\cite{guyon_jmlr03}, Correlation-based Feature Subset Selection
(CS)~\cite{Hall1998}, Reliff (RF)~\cite{Robnik-Sikonja2003} , Minimum
Redundancy Maximum Relevancem(MRMR)~\cite{peng2005} and our proposed
methods: Disjunction Voting Expectation (DVE) and Disjunction Voting
likelihood (DVL), Conjunction Voting Expectation (CVE), Conjunction
Voting likelihood (CVL).  The first five methods are ranking methods
that select features by ranking them with some metric and the last
seven are subset rank methods that assess subsets of features and
consider previous selections to decide the next selection.  The final
four methods were introduced and derived in the previous section.

@@

In our evaluation, we wish to answer the following questions:
Do different feature selection methods perform better with
each classifier or on each dataset in comparison to others?
How reliably do the different feature selection methods perform at
different stages of feature selection?  
Do any descriptive dataset features suggest where the different 
algorithms may perform well?

To answer these questions, we first start with
Figure~\ref{fig:perf_vs_dataset}, where we evaluate the performance of
various feature selection algorithms per dataset, averaged across
classifier type and each stage of feature selection. We can observe
here that our approaches generally perform well on recall (at a
reasonable accuracy level).  Notably, CVL performs best and
significantly better than competing methods on the NG and S datasets,
which we remark have the highest feature-to-data ratio and hence 
represent noisy cases for feature selection.  
%These results 
%indicate that the strong conjunctive voting scheme of CVL may help reduce 
%noise in feature selection for these datasets as previously conjectured.

% TODO Scott

In Figure~\ref{fig:perf_vs_classifier}, we examine how well our 
novel feature selection algorithms DVE, DVL, CVE, CVL 
perform vs. classifier, noting that overall performance is best
for Naive Bayes followed by Logistic Regression and noticeably worse for
the SVM Linear --- it seems that the probabilistic nature of our feature selection
dovetail best with classifiers that are themselves probabilistic,
and overall best with Naive Bayes which makes the same independence assumptions.

Next we examine Figure~\ref{fig:perf_vs_fs_alg}, where we evaluate the
performance distribution of various feature selection algorithms for
two datasets -- NG and S -- and classifier types -- Logistic
Regression and Naive Bayes -- at each stage of feature selection.
This graph shows that our CVL and CVE high recall feature selection
approaches provide strong recall performance at all stages of
feature selection in comparison to other selection methods.
%whose performance varies widely.
%We conjecture the reason for this stems from the strong agreement of
%features provided by the conjunctive voting scheme that helps reduce noise
%during feature selection process.

### 2014-02-05  feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/intro.tex
Latest submitted version

Feature selection~\cite{guyon_jmlr03}

While historically, feature selection

Naturally do not want to optimize for precision and recall directly (all false trivially optimizes precision, while all true trivially optimizes accuracy).  But when selecting features for a classifier intended to maximize accuracy, we may have a secondary preference for features that improve precision (reduce false positive rate) or improve recall (reduce false negative rate).  To this end, we remark that no previous work seems to have examined feature selection in the context of the precision-recall trade-off/

% Part of the reason is that previous feature selection has not been specific to classification whereas precision/recall are

@@

in well-known contingency table

\begin{center}
\begin{tabular}{l|l|l|} 
\multicolumn{1}{l}{} &  \multicolumn{1}{l}{Label T} & \multicolumn{1}{l}{Label F} \\ \cline{2-3}
Predicted T & TP & FP \\ \cline{2-3}
Predicted F & FN & TN \\ \cline{2-3}
%Predicted T & TP & FP & \#$T_P$ \\ \cline{2-3}
%Predicted F & FN & TN & \#$F_P$ \\ \cline{2-3}
%          &   & \#$T_A$ & \#$F_A$ & \\ \cline{2-3}
\end{tabular}
\end{center}

accuracy = tp + tn / tp + fp + fn + tn

precision = tp / tp + fp

recall = tp / tp + fn

can switch in columns, but typically predicting more false hurts recall and more true hurts precision

So precision and recall become secondary objectives after accuracy... provides flexibility in feature selection that has not been explicitly considered.

@@

\COMMENT
key insight: literature suggests redundancy is bad idea, but in fact, it is critical for precision (implies agreement)

>>>>

Feature selection is important in many practical supervised learning
settings to address computational constraints (e.g., in large-scale or
online learning) or to prevent overfitting when the data may be
high-dimensional but contain relatively few samples (e.g., as may
occur in medical or bioinformatics domains where features are abundant
but data is costly to obtain)~\cite{guyon_jmlr03}.

While a wide variety of feature selection methods have been proposed
in the literature, little research seems to focus on feature selection
specifically targeted to improve recall (minimization of false
negatives) in binary classification.  One reason for this is simply
that most feature selection methods are agnostic to the particular
supervised learning task --- applying to tasks from classification to
regression --- and hence are not focused on performance properties
specific to binary classifiers.  Yet, recall is an important aspect of
many binary classification problems (e.g., minimizing false negatives
in the identification of cancerous tumors) and it is critical to have
feature selection algorithms that can target features which facilitate
high-recall classification.

Naturally though, it would not make sense for a classifier or feature
selection method to focus on recall alone since the classifier which
always predicts \emph{true} (independent of the data) obtains optimal
recall performance.  To trade off precision with recall in order to
maintain high accuracy, one often uses a geometric average of the two
instead (known as F-score) but in practice most classifiers directly
optimize accuracy or some surrogate, e.g., a convex surrogate of 0-1
loss as in the SVM or maximum (conditional) likelihood as in Naive
Bayes or logistic regression.  So how can one achieve a high recall
SVM, Naive Bayes, or logistic regression classifier that already has a
well-defined accuracy-focused optimization criterion?  The idea we
pursue in this paper is that feature selection can help modulate the
recall performance of existing classifiers by encouraging selection of
features that cover more of the true cases in the data (hence
discouraging false negatives harmful to recall).

In this paper we propose a novel probabilistic voting model of feature
selection with the intent of encouraging false negative reduction
while still focusing on accuracy.  We argue that choosing features so
as to maximize likelihood objectives w.r.t. this model provides an
effective method for high recall feature selection.  Specifically, for
each objective, we derive an efficient formula for feature selection
in the filtering framework (i.e., greedy forward selection) and we
empirically compare the resulting feature selection criteria to a wide
variety of existing filtering methods.

Our results demonstrate that our high recall feature selection
approach does indeed improve recall with improvements most significant when 
there is a high feature to data ratio.  Such results provide a novel
and efficient feature selection algorithm to target high-recall
classification tasks.
% TODO: elaborate more

\COMMENT

@@

%can switch in columns, but typically predicting more false hurts recall and more true hurts precision

%So precision and recall become secondary objectives after accuracy... provides flexibility in feature selection that has not been explicitly considered.

@@

key insight: literature suggests redundancy is bad idea, but in fact, it is critical for precision (implies agreement)
